---
title: "technical report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


I. ABSTRACT


We created a web application, PicBook, that displays image representations of the most commonly tweeted nouns in any given location. By using API data from Twitter, Google Maps, and a word-to-icon generating site called The Noun Project, PicBook generates and displays pictorial icons depicting the most tweeted nouns in any user-entered region (address or coordinates/radius). The app, which is precoded with a list of 100 common nouns, sends the user-input location through the Google Places API, the Twitter API, and the Noun Project API to diplay the region's top 3 most commonly tweeted nouns (of the 100) and their corresponding icons. To exemplify the utility of our app, we built additional tabs into the Shiny interface, which display static graphics that show different temporal and spatial metrics for a set of precollected data.


II. INTRODUCTION


Pictorial writing is one of the oldest forms of written communication that dates farther back in history than spoken language. Conversely, social media is one of the most widely used forms of communication in the modern age. What can reconciling these two form of written language tell us about the relationships between words and pictures? How can we use pictorial representations to better understand the impact words and their usages? 

To examine this, we built a Shiny app, Picbook, that converts words used in Twitter posts into pictorial icons. Our app, precoded with a list of 100 common nouns, allows the user to type in a location, either an address or coordinates/radius; when entered, it outputs the top 3 of the 100 nouns most frequently used in the (approximately) 500 most recent tweets sent from that location. The application inputs the user-entered region through the Google Places API, which geographically identifies and generates a geocode for that region. The geocode is then input through the Twitter API, which identifies the most recent tweets (a specified number) published within that geographical region. Nouns are then parsed out from those tweets; of the 100 pre-coded nouns, the three most frequently used are looked up in the The Noun Project’s API data, which finally generates and displays the icons corresponding to those nouns. The app also includes 3 tabs that display different temporal and spatial metrics: the proportion of usage of a selected word across states, the usage frequency of a selected word across different times in a day, and the top most used word in major cities worldwide. 

Through this app and the visuals we produced through it, we aimed to accomplish 3 goals: to create a dynamic tool that can answer a variance of questions about most tweeted words across regions; to create an evocative type of output that presents data for a popular social tool in an engaging way; and to examine the way in which visuals can increase or change the impact of results. 


III. DATA


First, the following packages must be installed and libraries loaded.
```{r}
install.packages("httr")
install.packages("rtweet")
install.packages('maps')

library(httr)
library(rtweet)
library(readr)
library(mdsr)
library(tidyr)
library(maps)
```

In order to run this code properly, several API keys and tokens are required. Make sure you have access to the Google Places API (https://developers.google.com/places/web-service/intro), the standard Twitter APIs through a developer account (https://developer.twitter.com/en/docs/basics/getting-started), and the Noun Project api through a Noun Project account (https://api.thenounproject.com/getting_started.html). Then, make a file called `config.R` with the following variables saved with the appropriate values (as strings):
```{r}
 google_key <- '*Insert Google Places api key here*'
 twitter_key <- '*Insert Twitter api public key here*'
 twitter_secret_key <- '*Insert Twitter api secret key here*'
 twitter_token <- '*Insert Twitter api public token here*'
 twitter_secret_token <- '*Insert Twitter api secret token here*'
 nouns_key <- '*Insert nounproject api public key here*'
 nouns_secret_key <- '*Insert nounproject api secret key here*'
```

Then, one can load these values locally.
```{r}
source('config.R')
```

In order to generate geographic boundaries to search for tweets in we made use of the Google Places API. Using the `lookup_coords' function from `rtweet` one can search with some address criterion and get a boundary defined by coordinates.
```{r}
address <- '#Enter address here'
q <- lookup_coords(address, apikey = google_key)
```

One can also manually generate a search region defined by a single point (and its latitude and longitude) and a radius.
```{r}
lat <- '#Latitude number here'
long <- '#Longitude number here'
radius <- '#Radius number here'
q <- paste(lat, long, paste(radius, "km", sep = ""), sep = ",")
```

We gather our tweet data from the Twitter API with the `search_tweets` function from the `rtweet` package. First we define a token with all the appropriate key and token values (defined in `config.R`), then we run the `search_tweets` function for our geocode of `q` defined above. The result is a df with many columns but we are only concerned with the `text` column that contains the actual content of each tweet from our sample. The `search_tweets` function returns a number of recent tweets for the given geographic area. If the geographic are is defined to be small, the returned number of tweets may be smaller than the defined maximum number of tweets to be searched (`tweet_n`).
```{r}
create_token(app = "PicBook",
             consumer_key = twitter_key,
             consumer_secret = twitter_secret_key,
             access_token = twitter_token,
             access_secret = twitter_secret_token)

# Runs search of recent tweets
tweet_n <- 500 #Default, can change
rt <- search_tweets(
 "lang:en", geocode = q, n = tweet_n)
```

We've compiled a list of 100 common English nouns from https://www.espressoenglish.net/100-common-nouns-in-english/. The file consists of a single column, Words, with each word listed. Other files of nouns (or other words) saved in a similar formate could also be used instead, we simply chose a relatively short list in the interest of processing time.
```{r}
words_file <- read_csv('words.csv')
# Converts the words columns into a list for later reference
words <- words_file$Words
```

We also make use of the Noun Project API. This has been done using the `rtweet` package. First we define an oauth_app (a data type that is essentially a public and private key packaged together) and then define a function (taken from https://community.rstudio.com/t/getting-httr-to-work-with-the-noun-project-api/16165/2) to run GET requests through the API. We use GET requests with an endpoint to search for icons with a given name. This GET request returns a large amount of metadata but we only use a link to a url displaying the returned icon contained within this result.
```{r}
nouns_app <- oauth_app("nouns_api", nouns_key, nouns_secret_key)

get_nouns_api <- function(endpoint, baseurl = "http://api.thenounproject.com/", app) {
  url <- modify_url(baseurl, path = endpoint)
  info <- oauth_signature(url, app = app)
  header_oauth <- oauth_header(info)
  GET(url, header_oauth)
} 
```

Using the tweet data obtained from the Twitter API, we then used the words list loaded earlier to search for matches between words in the list and words in the tweets from our tweets df. This was done for each tweet in the df. Then, we added up the number of matches for each word over the entire sample of tweets. From there, we defined two functions: one that defines a list with the top n words by number of matches and one that defines a list with urls leading to .pngs with icons (from Noun Project API) that represent the top n matches. These urls are generated, for each top word, using the `get_nouns_api` function defined earlier, extracting the `preview_url` value from the result.
```{r}
match_words <- function(tweet){
  matched_values <- sapply(words, grepl, tweet)
  return(matched_values)
}

matched_words <- data.frame(mapply(match_words, rt$text))  %>%
  cbind(rownames(text_matches)) %>%
  rename(word = 'rownames(text_matches)') %>%
  mutate(occurences = rowSums(. == TRUE)) %>%
  select(word, occurences) %>%
  arrange(desc(occurences)) %>%
  select(word)

top_n_words <- function(words, n) {
  # Defines list with the n most common words
  top_words <- list()
  for (i in 1:n) {
    top_words[i] <- as.character(words$word)[i]
  }
  return(top_words)
}

top_n_urls <- function(words, n){
  # Generates endpoints for words from top_n_words function and then queries a response from nounproject API
  # A url with a link to a .png of the corresponding icon is then taken from this response
  urls <- list()
  for (i in 1:n) {
    endpoint <- paste("icon", as.character(words$word)[i], sep = "/")
    res <- get_nouns_api(endpoint = endpoint, app = nouns_app)
    icon_res <- content(res,"parsed")
    urls[i] <- icon_res$icon$preview_url
  }
  
  return(urls)
}
```

In addition to our dynamic, interactive generation of icons for the top tweeted words we also prepared static data on the usage of the same words in tweets. 
```{r}
state_data <- read_csv('state_data.csv')

state_data <- data.frame(words) %>%
  rename(word = words)
for (name in state.name) {
  q = lookup_coords(name, apikey = google_key)

  rt <- search_tweets(
  "lang:en", geocode = q, n = 250)
  
  num_rows <- rt %>%
    summarise(n())
  num_rows$`n()`
  
  rt_words <- most_common_words(rt) %>%
    mutate(occurences = occurences / num_rows$`n()`) %>%
    inner_join(state_data, by = 'word') %>%
    rename(!!name := occurences)
  state_data <- rt_words
  
}
```


```{r}
time_data_read <- read_csv('time_data.csv')

time_data <- data.frame(words) %>%
  rename(word = words)

num_rows <- rt_19 %>%
    summarise(n())
  num_rows$`n()`
  
rt_words <- most_common_words(rt_19) %>%
  mutate(occurences = occurences / num_rows$`n()`) %>%
  inner_join(time_data, by = 'word') %>%
  rename('19' = occurences)
time_data <- rt_words

time_data_trans <- time_data %>%
  gather(Time, proportion, -word) %>%
  spread(word, proportion) %>%
  mutate(Time = as.numeric(Time))
```


IV. RESULTS


```{r}
total_noun_num <- 3 #Change to search for more icons
url_list <- top_n_urls(matched_words, total_noun_num)

current_noun_num <- 1 #Change to see other icons
icon <- image_read(url_list[current_noun_num])
print(icon)
```

```{r}
selected_word_state <- ''

ggplot(state_data, aes_string(x = 'long', y = 'lat', fill = selected_word_state, group = 'map_group'), color = "white") + 
    coord_fixed(1.3) +
    geom_polygon(alpha = 0.7) +
    geom_polygon(color = "black", fill = NA) +
    theme_bw() +
    theme(
      plot.title = element_text(hjust = 0.5),
      axis.text = element_blank(),
      axis.line = element_blank(),
      axis.ticks = element_blank(),
      panel.border = element_blank(),
      panel.grid = element_blank(),
      axis.title = element_blank()) +
    labs(title = "Average Usage of Words Per Tweet Per State (Data generated at 3:00PM EST 12/10/18)") +
    scale_fill_gradient("Average usage")
```

```{r}
selected_word_times <- ''

ggplot(time_data_read, aes_string(x='Time', y=selected_word_time)) + geom_line() + geom_point() +
       labs(title = "Average Usage of Words Per Tweet Per Hour (Data generated for California from 6:00AM - 7:00PM PST 12/10/18)") +
       scale_x_continuous("Hour (24 hr format)", breaks = c(6,7,8,9,10,11,12,13,14,15,16,17,18,19)) +
       scale_y_continuous("Average Word Usage Per Tweet")

```

```{r}

```

V. CONCLUSIONS





Unlike the interactive Shiny app, which collects live data, the information displayed in thee tabs is collected in advance and thus shows only a static representation of data for a specific date and time. One purpose of building this app is to examine what kinds of words are being tweeted and how this differs across regions, while another purpose is simply to provide an evocative visual representation of a popular mode of social communication.
